Temporal stability in human interaction networks: sector sizes, topological prominence and time activity

### 2015-08-21 paper.tex
incorporadas modificacoes do chu

%\title[Evolution of interaction networks]{On the evolution of interaction networks: primitive typology of vertex, prominence of measures and activity statistics}% Force line breaks with \\
%\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
%\title[Stability of interaction networks]{Stability in human interaction networks: sector relative sizes, prominence of topological measures and time activity statistics.}% Force line breaks with \\
%\title[Stability in human interaction networks]{Sector relative sizes and topological metrics time stability in human interaction networks}% Force line breaks with \\
\title[Stability of interaction networks]{Temporal stability in human interaction networks: sector sizes, topological prominence and time activity}% Force line breaks with \\
\author{Renato Fabbri}%
\homepage{http://ifsc.usp.br/~fabbri/}
\email{fabbri@usp.br}
\affiliation{ 
	S\~ao Carlos Institute of Physics, University of S\~ao Paulo (IFSC/USP)%\\This line break forced with \textbackslash\textbackslash
}
%
%\author{Vilson V. da Silva Jr.}
%\homepage{http://automata.cc/}
%\email{vilson@void.cc}
%\altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
%
\author{Ricardo Fabbri}
\homepage{http://www.lems.brown.edu/~rfabbri/}
\email{rfabbri@iprj.uerj.br}
\altaffiliation{
	Instituto Polit\'ecnico, Universidade Estadual do Rio de Janeiro (IPRJ)
}%Lines break automatically or can be forced with \\

@@

The average and the dispersion were taken using circular statistics, in which each $measurement$ (data point) is represented as a unit complex number, $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measurement\frac{2\pi}{T}$, where T is the period in which the counting is repeated. E.g. $\theta=12\frac{2\pi}{24}=\pi$ for a message sent at $12h$ and given $T=24h$ for days. The moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:

@@

\begin{itemize}
	\item Exclusivist criterion $C_1$:  vertices are only classified if the class is the same according to all metrics. In this case, vertices classified do not usually reach 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
	\item Inclusivist criterion $C_2$: a vertex has the class given by any of the metrics. Therefore, a vertex may belong to more than one class, and the total number of members may exceed 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
	\item Exclusivist cascade $C_3$: vertices are only classified as hubs if they are hubs according to all metrics. Intermediary are the vertices classified either as intermediary or hubs with respect to all metrics. The remaining vertices are regarded as peripheral.
	\item Inclusivist cascade $C_4$: vertices are hubs if they are classified as so according to any of the metrics. The remaining vertices are classified as intermediary if they belong to this category for any of the metrics. Peripheral vertices will then be those which were not classified as hub or intermediary with any of the metrics. 
	\item Exclusivist externals $C_5$: vertices are only hubs if they are classified as such according to all the metrics. The remaining vertices are classified as peripheral if they fall into the periphery or hub classes by any metric. The rest of the nodes are classified as intermediary.
	\item Inclusivist externals $C_6$: hubs are vertices classified as hubs according to any metric. The remaining vertices will be peripheral if they are classified as such according to any metric. The rest of the vertices will be intermediary vertices.
\end{itemize}

@@

\subsection{Principal Component Analysis of topological metrics}\label{sec:pca}
Principal Component Analysis (PCA) is a well documented technique~\cite{pca} and was used to know: 1) what metrics contribute to each principal component and in which proportion; 2) how much of the dispersion is concentrated in each component; 3) expected values and dispersions for these quantities over various networks.

Be $\mathbf{X}=\{X[i,j]\}$ a matrix of all vertices $i$ and the respective
values for each metric $j$, 
$\mu_X [j]=\frac{\sum_i X[i,j]}{I}$ the mean of metric $j$ over all $I$ vertices, 
$\sigma_X [j]=\sqrt{\frac{\sum_i (X[i,j]-\mu_X [j])^2}{I}}$ the standard deviation of metric $j$,
and $\mathbf{X'}=X'[i,j]={\frac{X[i,j]-\mu_X[j]}{\sigma_X[j]}}$ 
the matrix with the \emph{z-score} of each metric. 
Be $\mathbf{V}=\{V[j,k]\}$ the matrix $J\times J$ of eigenvectors
of the covariance matrix $\mathbf{C}$
of $\mathbf{X'}$, one eigenvector per column.
Each eigenvector combines the original measures into one principal component, therefore
$V'[j,k]=100*\frac{|V[j,k]|}{\sum_{j'} |V[j',k]|}$
delivers the percentage of the principal component $k$
that is due to the measure $j$.
With $k$ eigenvectors 
$D[k]$,
it is enough to observe $D'[k]=100*\frac{D[k]}{\sum_{k'}D[k']}$
to know the percentage of dispersion that each principal component
is responsible for.
With the eigenvalues $k$ in ascending order
we observe, in general, the three first eigenvalues and
the respective eigenvectors in percentages:
$\{(V'[j,k],\;D'[k])\}$.
These usually reveal sufficient patterns for a sound analysis and sum between 60 and 95\% of the dispersion.
Namely, given $L$ observations $l$, each with $k$ pairs of
eigenvalues and eigenvectors, we observe
the mean
$\mu_{V'}[j,k]$
and the standard deviation $\sigma_{V'}[j,k]$ 
of the metric $j$ in the principal component $k$,
and the mean
$\mu_{D'}[k]$
and the standard deviation 
$\sigma_{D'}[k]$
of the contribution of the component $k$ to the dispersion
of the system:

>>>>

%\title[Evolution of interaction networks]{On the evolution of interaction networks: primitive typology of vertex, prominence of measures and activity statistics}% Force line breaks with \\
%\title[Evolution of interaction networks]{On the evolution of interaction networks: a primitive typology of vertex}% Force line breaks with \\
%\title[Stability of interaction networks]{Stability in human interaction networks: sector relative sizes, prominence of topological measures and time activity statistics.}% Force line breaks with \\
%\title[Stability in human interaction networks]{Sector relative sizes and topological metrics time stability in human interaction networks}% Force line breaks with \\
\title[Stability of interaction networks]{Temporal stability in human interaction networks: sector sizes, topological prominence and time activity}% Force line breaks with \\
\author{Renato Fabbri}%
\homepage{http://ifsc.usp.br/~fabbri/}
\email{fabbri@usp.br}
\affiliation{ 
	S\~ao Carlos Institute of Physics, University of S\~ao Paulo (IFSC/USP),
	PO Box 369, 13560-970, S\~ao Carlos, SP, Brazil %\\This line break forced with \textbackslash\textbackslash
}
%
%\author{Vilson V. da Silva Jr.}
%\homepage{http://automata.cc/}
%\email{vilson@void.cc}
%\altaffiliation[Also at ]{IFSC-USP}%Lines break automatically or can be forced with \\
%
\author{Ricardo Fabbri}
\homepage{http://www.lems.brown.edu/~rfabbri/}
\email{rfabbri@iprj.uerj.br}
\altaffiliation{
	Instituto Polit\'ecnico, Universidade Estadual do Rio de Janeiro (IPRJ)
}%Lines break automatically or can be forced with \\

@@

The average and the dispersion were taken using circular statistics, in which each $measurement$ (data point) is represented as a unity complex number, $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measurement\frac{2\pi}{T}$, T is the period in which the counting is repeated. For example, $\theta=12\frac{2\pi}{24}=\pi$ for a message sent at $12h$ and given $T=24h$ for days. The moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:

@@

\begin{itemize}
	\item Exclusivist criterion $C_1$:  vertices are only classified if the class is the same according to all metrics. In this case, vertices classified do not usually reach 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
	\item Inclusivist criterion $C_2$: a vertex has the class given by any of the metrics. Therefore, a vertex may belong to more than one class, and the total number of members may exceed 100\%, which is indicated by a black line in Figure~\ref{fig:sectIL}.
	\item Exclusivist cascade $C_3$: vertices are only classified as hubs if they are hubs according to all metrics. Intermediary are the vertices classified either as intermediary or hubs with respect to all metrics. The remaining vertices are regarded as peripheral.
	\item Inclusivist cascade $C_4$: vertices are hubs if they are classified as such according to any of the metrics. The remaining vertices are classified as intermediary if they belong to this category for any of the metrics. Peripheral vertices will then be those which were not classified as hub or intermediary with any of the metrics. 
	\item Exclusivist externals $C_5$: vertices are only hubs if they are classified as such according to all the metrics. The remaining vertices are classified as peripheral if they fall into the periphery or hub classes by any metric. The rest of the nodes are classified as intermediary.
	\item Inclusivist externals $C_6$: hubs are vertices classified as hubs according to any metric. The remaining vertices will be peripheral if they are classified as such according to any metric. The rest of the vertices will be intermediary vertices.
\end{itemize}

@@

\subsection{Principal Component Analysis of topological metrics}\label{sec:pca}
Principal Component Analysis (PCA) is a well documented technique~\cite{pca} and was used to learn: 1) which metrics contribute to each principal component and in which proportion; 2) how much of the dispersion is concentrated in each component; 3) expected values and dispersions for these quantities over various networks.

Let $\mathbf{X}=\{X[i,j]\}$ be a matrix of all vertices $i$ and the respective
values for each metric $j$, 
$\mu_X [j]=\frac{\sum_i X[i,j]}{I}$ the mean of metric $j$ over all $I$ vertices, 
$\sigma_X [j]=\sqrt{\frac{\sum_i (X[i,j]-\mu_X [j])^2}{I}}$ the standard deviation of metric $j$,
and $\mathbf{X'}=X'[i,j]={\frac{X[i,j]-\mu_X[j]}{\sigma_X[j]}}$ 
the matrix with the \emph{z-score} of each metric. 
Let $\mathbf{V}=\{V[j,k]\}$ be the matrix $J\times J$ of eigenvectors
of the covariance matrix $\mathbf{C}$
of $\mathbf{X'}$, one eigenvector per column.
Each eigenvector combines the original measures into one principal component, therefore
$V'[j,k]=100*\frac{|V[j,k]|}{\sum_{j'} |V[j',k]|}$
delivers the percentage of the principal component $k$
that is due to the measure $j$.
With $k$ eigenvectors 
$D[k]$,
it is enough to observe $D'[k]=100*\frac{D[k]}{\sum_{k'}D[k']}$
to know the percentage of dispersion that each principal component
is responsible for.
With the eigenvalues $k$ in ascending order
we observe, in general, the three first eigenvalues and
the respective eigenvectors in percentages:
$\{(V'[j,k],\;D'[k])\}$.
These usually reveal sufficient patterns for a sound analysis and sum between 60 and 95\% of the dispersion.
Namely, given $L$ observations $l$, each with $k$ pairs of
eigenvalues and eigenvectors, we observe
the mean
$\mu_{V'}[j,k]$
and the standard deviation $\sigma_{V'}[j,k]$ 
of the metric $j$ in the principal component $k$,
and the mean
$\mu_{D'}[k]$
and the standard deviation 
$\sigma_{D'}[k]$
of the contribution of the component $k$ to the dispersion
of the system:
